\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-07-21 paper.qmd
Update the figure title of MNIST PaCMAP layout

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = b, 
                         y = log(MSE),
                         color = buff,
                         group = buff)) + 
  geom_point(size = 0.05) +
  geom_line(linewidth = 0.05) + 
  geom_vline(xintercept = 220, linetype="solid", 
             color = "black", linewidth=0.1) +
  geom_vline(xintercept = 112, linetype=2, 
             color = "black", linewidth=0.1) +
  geom_vline(xintercept = 364, linetype=2, 
             color = "black", linewidth=0.1) +
  scale_colour_continuous_sequential(palette = "Magenta") +
  labs(x = expression(b), y = "log(MSE)", color = expression(q)) +
  ggtitle("(a)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 5),
        axis.title.x = element_text(size = 5),
        axis.title.y = element_text(size = 5, 
            angle=90),
        title = element_text(size = 72),
        legend.title = element_text(size = 5),
        legend.text = element_text(size = 3))
```

@@

Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>).
:::

@@

<!--tSNE with perplexity:30-->
```{r}
#| echo: false
#| fig-cap: "$2\\text{-}D$ layout from tSNE applied for the PBMC3k dataset.  Is this a best representation of the original data? The parameter setting is perplexity: 30."
#| label: fig-tsne-suggest
#| fig-pos: H
#| out-width: 40%

@@

<!--bin1 = 15, bin2 = 28, b = 420, non_empty = 136-->
```{r}
#| echo: false
#| fig-cap: "(a) Model generated in $2\\text{-}D$ with tSNE, and (b) $p\\text{-}D$ model error in $2\\text{-}D$. The $2\\text{-}D$ model shows three well-separated distant clusters. The $p\\text{-}D$ model errors are distributed along clusters, but most low $p\\text{-}D$ model errors present in the large cluster."
#| label: fig-model-pbmc
#| fig-pos: H
#| out-height: 30%

@@

Model in \gD{}, on the tSNE layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/5Y1hE4i7N2k>).
:::

@@

<!--PaCMAP param: n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0-->
```{r}
#| fig-cap: "$2\\text{-}D$ layout from PaCMAP applied for the digit 1 of the MNIST dataset with a specific parameter settings. The angle of the digit 1 varies along this structure. Images at the top-left of the $2\\text{-}D$ layout show the digit 1 angled more to the left, while images at the bottom-right show the digit 1 angled more to the right. Is this a reasonable representation of the digit 1?"
#| label: fig-pacmap-author
#| fig-pos: H
#| out-height: 22%

@@

Model in \gD{}, on the PaCMAP layout, and three views of the fit in projections from $10\text{-}D$, for the digit 1 of MNIST data. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/zcg_GXBmqjA>).

>>>>

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = b, 
                         y = log(MSE),
                         color = buff,
                         group = buff)) + 
  geom_point(size = 0.05) +
  geom_line(linewidth = 0.05) + 
  geom_vline(xintercept = 220, linetype="solid", 
             color = "black", linewidth=0.1) +
  geom_vline(xintercept = 112, linetype=2, 
             color = "black", linewidth=0.1) +
  geom_vline(xintercept = 364, linetype=2, 
             color = "black", linewidth=0.1) +
  scale_colour_continuous_sequential(palette = "Magenta") +
  labs(x = expression(b), y = "log(MSE)", color = expression(q)) +
  ggtitle("(a)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7, 
            angle=90),
        title = element_text(size = 72),
        legend.title = element_text(size = 7),
        legend.text = element_text(size = 5))
```

@@

Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ 0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$). (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>).
:::

@@

<!--tSNE with perplexity:30-->
```{r}
#| echo: false
#| fig-cap: "The $2\\text{-}D$ layout from tSNE applied for the PBMC3k dataset. The default (hyper-)parameter setting is a perplexity $30$. We use our model-in-the-data-space to assess whether this is an accurate representation of structure present in the high-dimensional data, or if it is misleading."
#| label: fig-tsne-suggest
#| fig-pos: H
#| out-width: 40%

@@

<!--bin1 = 15, bin2 = 20, b = 300, non_empty = 136-->
```{r}
#| echo: false
#| fig-cap: "(a) Model generated in $2\\text{-}D$ with tSNE, and (b) $p\\text{-}D$ model error in $2\\text{-}D$. The $2\\text{-}D$ model shows three well-separated distant clusters. The $p\\text{-}D$ model errors are distributed along clusters, but most low $p\\text{-}D$ model errors present in the large cluster."
#| label: fig-model-pbmc
#| fig-pos: H
#| out-height: 30%

@@

Model in \gD{}, on the tSNE layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.058)$, $b = 300 \  (15, \ 20)$, $m = 136$, benchmark value to remove large edges is $0.133$). (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/5Y1hE4i7N2k>).
:::

@@

<!--PaCMAP param: n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0-->
```{r}
#| fig-cap: "$2\\text{-}D$ layout from PaCMAP applied for the digit 1 of the MNIST dataset. The (hyper-)parameter settings, beyond the defaults, are 10 nearest neighbors, ratio of the number of mid-near pairs to the number of neighbors equal to 0.5, and the ratio of the number of further pairs to the number of neighbors is 2. We use our model-in-the-data-space to assess whether this is an accurate representation of structure present in the high-dimensional data, or if it is misleading. The angle of the digit 1 varies along this structure. Images at the top-left of the $2\\text{-}D$ layout show the digit 1 angled more to the left, while images at the bottom-right show the digit 1 angled more to the right."
#| label: fig-pacmap-author
#| fig-pos: H
#| out-height: 22%

@@

Model in \gD{}, on the PaCMAP layout, and three views of the fit in projections from $10\text{-}D$, for the digit 1 of MNIST data ($(s_1, \ s_2) = (-0.100, \ -0.059)$, $b = 374 \  (22, \ 17)$, $m = 140$, benchmark value to remove large edges is $0.094$). (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/zcg_GXBmqjA>).

### 2024-07-21 paper.tex
Update the figure title of MNIST PaCMAP layout

\caption{\label{fig-pbmc1-sc}Model in \gD{}, on the UMAP layout, and
three views of the fit in projections from \(9\text{-}D\), for the
PBMC3k data. (The \textbf{langevitour} software is used to view the data
with a tour, and the full video is available at
\url{https://youtu.be/VqqWuE0Jj6A}).}

@@

\caption{\label{fig-tsne-suggest}\(2\text{-}D\) layout from tSNE applied
for the PBMC3k dataset. Is this a best representation of the original
data? The parameter setting is perplexity: 30.}

@@

\caption{\label{fig-pbmc2-sc}Model in \gD{}, on the tSNE layout, and
three views of the fit in projections from \(9\text{-}D\), for the
PBMC3k data. (The \textbf{langevitour} software is used to view the data
with a tour, and the full video is available at
\url{https://youtu.be/5Y1hE4i7N2k}).}

@@

\caption{\label{fig-pacmap-author}\(2\text{-}D\) layout from PaCMAP
applied for the digit 1 of the MNIST dataset with a specific parameter
settings. The angle of the digit 1 varies along this structure. Images
at the top-left of the \(2\text{-}D\) layout show the digit 1 angled
more to the left, while images at the bottom-right show the digit 1
angled more to the right. Is this a reasonable representation of the
digit 1?}

@@

\caption{\label{fig-mnist1-sc}Model in \gD{}, on the PaCMAP layout, and
three views of the fit in projections from \(10\text{-}D\), for the
digit 1 of MNIST data. (The \textbf{langevitour} software is used to
view the data with a tour, and the full video is available at
\url{https://youtu.be/zcg_GXBmqjA}).}

>>>>

\caption{\label{fig-pbmc1-sc}Model in \gD{}, on the UMAP layout, and
three views of the fit in projections from \(9\text{-}D\), for the
PBMC3k data (\((s_1, \ s_2) = (-0.050, \ 0.041)\),
\(b = 870 \  (30, \ 29)\), \(m = 135\), benchmark value to remove large
edges is \(0.099\)). (The \textbf{langevitour} software is used to view
the data with a tour, and the full video is available at
\url{https://youtu.be/VqqWuE0Jj6A}).}

@@

\caption{\label{fig-tsne-suggest}The \(2\text{-}D\) layout from tSNE
applied for the PBMC3k dataset. The default (hyper-)parameter setting is
a perplexity \(30\). We use our model-in-the-data-space to assess
whether this is an accurate representation of structure present in the
high-dimensional data, or if it is misleading.}

@@

\caption{\label{fig-pbmc2-sc}Model in \gD{}, on the tSNE layout, and
three views of the fit in projections from \(9\text{-}D\), for the
PBMC3k data (\((s_1, \ s_2) = (-0.050, \ -0.058)\),
\(b = 300 \  (15, \ 20)\), \(m = 136\), benchmark value to remove large
edges is \(0.133\)). (The \textbf{langevitour} software is used to view
the data with a tour, and the full video is available at
\url{https://youtu.be/5Y1hE4i7N2k}).}

@@

\caption{\label{fig-pacmap-author}\(2\text{-}D\) layout from PaCMAP
applied for the digit 1 of the MNIST dataset. The (hyper-)parameter
settings, beyond the defaults, are 10 nearest neighbors, ratio of the
number of mid-near pairs to the number of neighbors equal to 0.5, and
the ratio of the number of further pairs to the number of neighbors is
2. We use our model-in-the-data-space to assess whether this is an
accurate representation of structure present in the high-dimensional
data, or if it is misleading. The angle of the digit 1 varies along this
structure. Images at the top-left of the \(2\text{-}D\) layout show the
digit 1 angled more to the left, while images at the bottom-right show
the digit 1 angled more to the right.}

@@

\caption{\label{fig-mnist1-sc}Model in \gD{}, on the PaCMAP layout, and
three views of the fit in projections from \(10\text{-}D\), for the
digit 1 of MNIST data (\((s_1, \ s_2) = (-0.100, \ -0.059)\),
\(b = 374 \  (22, \ 17)\), \(m = 140\), benchmark value to remove large
edges is \(0.094\)). (The \textbf{langevitour} software is used to view
the data with a tour, and the full video is available at
\url{https://youtu.be/zcg_GXBmqjA}).}
